### 2024-09-27 Article/Article.tex
Changed layout

\documentclass[conference]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage{ifthen, xkeyval, tikz, calc}

\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
% \usepackage{svg}
\usepackage{subcaption}
\usepackage{float}
\usepackage{tikz}
\usepackage{lipsum}
\usepackage{booktabs}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\checkmark}%
\newcommand{\xmark}{\ding{55}}%

\graphicspath{{./figures/}}
%\graphicspath{{./tmp_figures/}}

% \let\marginpar\oldmarginpar
\setlength{\marginparwidth}{5cm}
% \newcommand{\Bnote}[1]{\textbf{[Boaz: #1]}}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[colorinlistoftodos,prependcaption, textwidth=3cm]{todonotes}

\title{
Statistical Theory\\
Chess Dataset Analysis
}

% Authors must not appear in the submitted version. They should be hidden
% as long as the \iclrfinalcopy macro remains commented out below.
% Non-anonymous submissions will be rejected without review.

\author{
   \IEEEauthorblockN{Dor Boker, Itamar Nakar}
   \IEEEauthorblockA{
      I.D: , 325829000\\
      Email: , itamar.nakar@gmail.com
   }
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\listfiles
\begin{document}

\maketitle

\section{Introduction}
The length of a chess match can be influenced by a variety of factors, including player skill level, strategy, and in-game dynamics. This study investigates the relationship between player ratings and the duration of chess matches, seeking to determine whether higher-rated players tend to play shorter or longer games. In addition to player ratings, we explore other game-related variables that might impact match length, such as opening moves and the type of result (win, draw, or loss).

Using a dataset of chess games, we apply statistical methods such as correlation analysis and regression models to analyze the influence of these factors. The study aims to offer insights into how skill level and game dynamics affect the length of a match, contributing to a broader understanding of player performance in chess.

\end{document}

### 2024-09-27 Article/Article.tex
Updated Article

@@
 \section{Introduction}
+Chess is one of humanity's oldest board games.
+It is played by two players, nicknamed White and Black over an $8\times8$ grid, where each players controls 16 pieces which correspond to its own color.
+The Goal of each player is to capture the opponents King piece(or, more accurately, to make it so the opponent's King cannot escape capture).
+
 The length of a chess match can be influenced by a variety of factors, including player skill level, strategy, and in-game dynamics.
-This study investigates the relationship between player ratings and the duration of chess matches, seeking to determine whether higher-rated players tend to play shorter or longer games.
-In addition to player ratings, we explore other game-related variables that might impact match length, such as opening moves and the type of result (win, draw, or loss).
+In this project we investigates the relationship between player ratings and the duration of chess matches, seeking to determine whether higher-rated players tend to play shorter or longer games.
+In addition to player ratings, we explore other game-related variables that might impact match length, such as opening moves, the type of result (checkkmate, resignation, or draw) and more.
 

### 2024-09-27 Article/Article.tex
Updated Article

@@
 The Goal of each player is to capture the opponents King piece(or, more accurately, to make it so the opponent's King cannot escape capture).
+Chess doesn't involve any luck, no hidden information.
+It is determined by the players' knowlage of the game, their strategical and analytical capabilities.
 
@@
 In this project we investigates the relationship between player ratings and the duration of chess matches, seeking to determine whether higher-rated players tend to play shorter or longer games.
-In addition to player ratings, we explore other game-related variables that might impact match length, such as opening moves, the type of result (checkkmate, resignation, or draw) and more.
+In addition to player ratings, we explore other game-related variables that might impact match length, such as opening moves, the type of result (checkmate, resignation, or draw) and more.
 

### 2024-09-29 Article/Article.tex
updated article

--- 2024-09-27 Article/Article.tex
@@
 \section{Introduction}
-The length of a chess match can be influenced by a variety of factors, including player skill level, strategy, and in-game dynamics.
-This study investigates the relationship between player ratings and the duration of chess matches, seeking to determine whether higher-rated players tend to play shorter or longer games.
-In addition to player ratings, we explore other game-related variables that might impact match length, such as opening moves and the type of result (win, draw, or loss).
+Chess is one of humanity's oldest board games, played by two players on an $8\times8$ grid.
+Each player, controlling 16 pieces of their color, aims to checkmate the opponent's King, making it impossible for the King to escape capture.
+Unlike many games, chess does not involve luck or hidden information; the outcome is determined solely by the players' knowledge, strategy, and analytical skills.
 
-Using a dataset of chess games, we apply statistical methods such as correlation analysis and regression models to analyze the influence of these factors.
-The study aims to offer insights into how skill level and game dynamics affect the length of a match, contributing to a broader understanding of player performance in chess.
+The length of a chess match is influenced by several factors, including the players' skill levels, strategic choices, and in-game dynamics.
+In this project, we investigate the relationship between player ratings and the duration of chess matches, aiming to determine whether higher-rated players tend to play shorter or longer games.
+In addition to player ratings, we examine other variables that may impact match length, such as opening strategies, game outcomes (win, loss, draw), and time controls.
+
+Our dataset comprises 19,113 games from Lichess, an online chess platform, each described by 16 features.
+The most relevant features to our analysis are: \begin{itemize} \item \textbf{Turns}: A positive integer representing the length of the game, where one turn includes a move by both White and Black.
+\item \textbf{ELO Rating}: A system that quantifies player skill, widely used by FIDE and online chess platforms, which we use to assess its correlation with match length.
+\item \textbf{Time Controls}: Competitive games are categorized into four main time control classes: Bullet, Blitz, Rapid, and Classical, based on initial clock times and increments.
+\item \textbf{Winner and Victory Status}: These columns capture whether White or Black won and the method of victory (checkmate, resignation, draw).
+\end{itemize} Through statistical analysis, we aim to understand how these factors influence the duration of chess matches and to provide insights into the relationship between skill level and game dynamics.
 

### 2024-09-29 Article/Article.tex
updated article

@@
 \usepackage[table]{xcolor}
-\usepackage{ifthen, xkeyval, tikz, calc}
 
+\usepackage{tikz, calc}
+\usepackage{cite}
 \usepackage{hyperref}
@@
 \usepackage{float}
-\usepackage{tikz}
-\usepackage{lipsum}
 \usepackage{booktabs}
@@
 % \newcommand{\Bnote}[1]{\textbf{[Boaz: #1]}}
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
 \usepackage[colorinlistoftodos,prependcaption, textwidth=3cm]{todonotes}
@@
 \section{Introduction}
-Chess is one of humanity's oldest board games, played by two players on an $8\times8$ grid.
+Chess is one of humanity's oldest board games \cite{chesswiki}, played by two players on an $8\times8$ grid.
 Each player, controlling 16 pieces of their color, aims to checkmate the opponent's King, making it impossible for the King to escape capture.
@@
 
+\section{Results}
+
+\newpage
+\bibliographystyle{IEEEtran}
+\bibliography{references}
+
 \end{document}

### 2024-09-29 Article/Article.tex
updated article
Updated Article

@@
    \IEEEauthorblockA{
-      I.D: , 325829000\\
-      Email: , itamar.nakar@gmail.com
+      I.D: 209271279 , 325829000\\
+      Email: dorboker@gmail.com, itamar.nakar@gmail.com
    }
@@
 \section{Results}
+\subsection{Data Transformation}
+
+To optimize the quality of our analysis, we undertook a cleaning process to remove entries that could distort or obscure meaningful insights from the dataset. Below are the key transformations:
+
+\begin{itemize}
+    \item \textbf{Uncompetitive Games}: Games marked as casual or non-competitive were excluded.
+Such games may not reflect players' full effort or strategic depth, and are thus not representative of a standard chess match.
+
+    \item \textbf{Resignations}: Games concluded by resignation were removed.
+A resignation introduces a psychological factor into game length and outcome, which can skew the statistical distribution.
+By removing these games, we aim to focus on games decided by normal play.
+
+    \item \textbf{Elo Adjustments}: New players are initially assigned an Elo rating of 1500, which can skew the distribution of player strength.
+To mitigate this, each player's rating was updated to reflect their most recent Elo score within the dataset.
+\end{itemize}
+
+After these trasformations, we are left with 5,827 games which we will now begin to analyze.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{gamma_fit.png}
+    \caption{Gamma distribution fitted to the 'Turns' data}
+    \label{fig:gamma_fit}
+\end{figure}
+\subsection{Distribution of 'Turns'}
+
+To better understand the distribution of the 'Turns' variable, we attempted to fit it to several probability distribution families.
+The Gamma distribution was selected first due to its ability to model skewed data.
+However, as shown in Figure \ref{fig:gamma_fit}, the Gamma distribution is too broad for this dataset, suggesting it does not fully capture the underlying structure of 'Turns.'
+
+To refine the fit, we combined the Gamma distribution with a Poisson distribution, which has a narrower shape.
+This combination results in a more accurate fit to the empirical data, as shown in Figure \ref{fig:gam_poi_fit}.
+The Gamma-Poisson mixture balances the excess spread of the Gamma distribution with the narrowness of the Poisson, providing a better representation of the 'Turns' distribution.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{gam_poi_fit.png}
+    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
+    \label{fig:gam_poi_fit}
+\end{figure}
 

### 2024-09-29 Article/Article.tex
Updated Article

@@
 
-Our dataset comprises 19,113 games from Lichess, an online chess platform, each described by 16 features.
+Our dataset \cite{dataset} comprises 19,113 games from Lichess, an online chess platform, each described by 16 features.
 The most relevant features to our analysis are: \begin{itemize} \item \textbf{Turns}: A positive integer representing the length of the game, where one turn includes a move by both White and Black.
-\item \textbf{ELO Rating}: A system that quantifies player skill, widely used by FIDE and online chess platforms, which we use to assess its correlation with match length.
+\item \textbf{ELO Rating}: A system that quantifies player skill, widely used by FIDE and online chess platforms \cite{ELO}, which we use to assess its correlation with match length.
 \item \textbf{Time Controls}: Competitive games are categorized into four main time control classes: Bullet, Blitz, Rapid, and Classical, based on initial clock times and increments.
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{allgames.png}
+    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
+    \label{fig:all_games}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{ratedgames.png}
+    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
+    \label{fig:rated_games}
+\end{figure}
+
+\subsection{Rating Distribution}
+The distribution of player ratings in the dataset initially resembles a normal distribution with a peak around the mean.
+However, this peak diminishes when games involving unrated players are excluded, bringing the distribution closer to normality.
+This is illustrated in Figure~\ref{fig:all_games} and Figure~\ref{fig:rated_games}.
+Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the extremes.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{ratingqq.png}
+    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
+    \label{fig:rating_qq}
+\end{figure}
+
+\subsubsection{Kolmogorov-Smirnov (KS) Test for Normality}
+We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution.
+The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed.
+However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size.
+This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality.
+
+\subsection{Relations Between Features}
+Analyzing the dataset’s convolution matrix reveals linear relationships between certain features.
+For example, there is a strong correlation between the ratings of white and black players in a game, which aligns with the expectation that players of similar ratings tend to compete against each other.
+However, no significant linear relationship was observed between the winner and the ratings of either player.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{ratingecdf.png}
+    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
+    \label{fig:ecdf_plot}
+\end{figure}
+
+\subsection{Connection Between Player Rating and Game Length}
+To investigate whether player ratings are associated with game length, we divided players into two groups based on their ratings: those with ratings higher than 1500 (high-rated) and those with ratings lower than 1500 (low-rated).
+The threshold of 1500 was chosen because it represents the starting rating on the Lichess platform.
+The empirical cumulative distribution function (ECDF) of game lengths for both groups (Figure~\ref{fig:ecdf_plot}) shows that high-rated players generally play longer games.
+
+\subsubsection{Mann-Whitney U Test for Game Length}
+Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players.
+Our null hypothesis was that high-rated players do not play longer games than low-rated players on average.
+The test returned a p-value of \(2.7 \times 10^{-73}\), which allows us to reject the null hypothesis.
+Therefore, we conclude that player rating significantly influences game length, with higher-rated players tending to play longer games.
+
+\subsection{Summary of Results}
+The analysis demonstrates a clear statistical relationship between player rating and game length.
+The Mann-Whitney U test confirmed that higher-rated players play longer games.
+This finding supports the hypothesis that player rating is a factor in game duration, with the distribution of ratings further informing our understanding of game dynamics.
+
 \newpage

### 2024-09-29 main.py

@@ -3,15 +3,11 @@ import pandas as pd
 from scipy import stats
 import matplotlib.pyplot as plt
 import seaborn as sns
+from src.transform import transform
 
 # Read the dataset
 df = pd.read_csv('games.csv')
 
-# Transforming the dataset and removing unnecesary information 
-df = df.drop(columns=['id', 'opening_name'])
-df.drop(df[df['rated'] == False].index, inplace=True)
-df['created_at'] = pd.to_datetime(df['created_at'], unit='ms')
-df['created_at'] = df['created_at'].dt.tz_localize('UTC')
-df['last_move_at'] = df['created_at'].dt.year
-df = df.rename(columns={'last_move_at': 'year'})
-print(df.head())
+# Transforming the dataset and removing unnecesary information
+df = transform(df)
+print(df.head(20))

### 2024-09-29 src/transform.py

@@ -1,0 +1,42 @@
+import numpy as np
+import pandas as pd
+
+def transform(df):
+    df = df.drop(columns=['id', 'opening_name', 'opening_ply', 'moves', 'opening_eco'])
+    df.drop(df[df['rated'] == False].index, inplace=True)
+    df['created_at'] = pd.to_datetime(df['created_at'], unit='ms')
+    df['created_at'] = df['created_at'].dt.tz_localize('UTC')
+    df['last_move_at'] = df['created_at'].dt.year
+    df = df.rename(columns={'last_move_at': 'year'})
+    df = df.rename(columns={'created_at': 'time'})
+    df = df.rename(columns={'increment_code': 'time_control'})
+    df = df[df['victory_status'] != 'resign']
+    df = df[df['victory_status'] != 'outoftime']
+    df = df.reset_index()
+    df = df.drop(columns=['index'])
+    df = updateELO(df)
+    df = df.rename(columns={'white_id': 'ave_rating'})
+    df = df.rename(columns={'black_id': 'dif_rating'})
+
+    for index, row in df.iterrows():
+        df.at[index, 'ave_rating'] = ave(row['white_rating'], row['black_rating'])
+        df.at[index, 'dif_rating'] = dif(row['white_rating'], row['black_rating'])  
+    return df
+
+def updateELO(df):
+    updatedELO = {}
+    for index, row in df.iterrows():
+        if row['white_id'] not in updatedELO or row['time'] > updatedELO[row['white_id']][1]:
+            updatedELO[row['white_id']] = (row['white_rating'], row['time'])
+        if row['black_id'] not in updatedELO or row['time'] > updatedELO[row['black_id']][1]:
+            updatedELO[row['black_id']] = (row['black_rating'], row['time'])
+
+    df['white_rating'] = df.apply(lambda row: updatedELO[row['white_id']][0], axis=1)
+    df['black_rating'] = df.apply(lambda row: updatedELO[row['black_id']][0], axis=1)
+    return df
+
+def ave(white_rating, black_rating):
+    return (white_rating + black_rating) // 2
+
+def dif(white_rating, black_rating):
+    return (white_rating - black_rating)

### 2024-09-29 main.py
Updated main and src

@@ -1,13 +1,21 @@
+import os
 import numpy as np
 import pandas as pd
 from scipy import stats
 import matplotlib.pyplot as plt
 import seaborn as sns
 from src.transform import transform
+from src.trasformation_plot import tr_plot
+from src.save_plot import save_plot
+
+save_dir = os.path.abspath('figures')
 
 # Read the dataset
 df = pd.read_csv('games.csv')
 
 # Transforming the dataset and removing unnecesary information
-df = transform(df)
-print(df.head(20))
+df_tr = transform(df)
+
+fig = tr_plot(df, df_tr)
+
+save_plot(fig, 'transformation_plot', save_dir)

### 2024-09-29 src/save_plot.py

@@ -1,0 +1,11 @@
+import os
+import matplotlib.pyplot as plt
+
+def save_plot(fig, name, save_dir):
+    if not os.path.exists(save_dir):
+        os.makedirs(save_dir)
+
+    try:
+        plt.savefig(os.path.join(save_dir, name+'.png'))
+    except Exception as e:
+        print(f"Error saving figure: {e}")

### 2024-09-29 src/trasformation_plot.py

@@ -1,0 +1,41 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import scipy.stats as stats
+
+def tr_plot(df1, df2):
+    # Generate random data for four datasets
+    data1 = df1['white_rating']
+    data2 = df2['white_rating']
+    data3 = df1['turns']
+    data4 = df2['turns']
+
+    data = [data1, data2, data3, data4]
+
+    # Set up the figure and axes
+    fig, axs = plt.subplots(2, 2, figsize=(10, 10))  # 2x2 grid of plots
+
+    for i, ax in enumerate(axs.flatten()):
+        # Plot the histogram
+        if i % 2 == 1:
+            color = 'b'
+        else:
+            color = 'g'
+        ax.hist(data[i], bins=50, density=True, alpha=0.6, color=color)
+        
+        # Fit a normal distribution to the data
+        mu, std = stats.norm.fit(data[i])
+        
+        # Plot the PDF of the fitted normal distribution
+        xmin, xmax = ax.get_xlim()
+        x = np.linspace(xmin, xmax, 100)
+        p = stats.norm.pdf(x, mu, std)
+        ax.plot(x, p, 'k', linewidth=2)
+        
+        # Title with the fit parameters
+        title1 = 'player rating' if i < 2 else 'turns'
+        title2 = ' before transformation' if i % 2 == 0 else ' after transformation'
+        ax.set_title(title1+title2)
+
+    return fig
+

### 2024-09-29 Article/Article.tex
Updated Article

@@
 \item \textbf{ELO Rating}: A system that quantifies player skill, widely used by FIDE and online chess platforms \cite{ELO}, which we use to assess its correlation with match length.
-\item \textbf{Time Controls}: Competitive games are categorized into four main time control classes: Bullet, Blitz, Rapid, and Classical, based on initial clock times and increments.
+\item \textbf{Time Controls}: Competitive games are categorized into four main time control classes: Bullet, Blitz, Rapid, and Classical, based on initial clock times and increments \cite{time}.
 \item \textbf{Winner and Victory Status}: These columns capture whether White or Black won and the method of victory (checkmate, resignation, draw).
@@
 However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size.
-This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality.
+This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.
 

### 2024-09-29 Article/Article.tex
fixed nothing

@@
 
-    \item \textbf{Resignations}: Games concluded by resignation were removed.
-A resignation introduces a psychological factor into game length and outcome, which can skew the statistical distribution.
-By removing these games, we aim to focus on games decided by normal play.
-
     \item \textbf{Elo Adjustments}: New players are initially assigned an Elo rating of 1500, which can skew the distribution of player strength.
@@
 \end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{transformation_plot.png}
+    \caption{Transformation's effected on Rating and Turns}
+    \label{fig:transformation_plot}
+\end{figure}
 
@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\linewidth]{allgames.png}
-    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
-    \label{fig:all_games}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\linewidth]{ratedgames.png}
-    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
-    \label{fig:rated_games}
-\end{figure}
-
 \subsection{Rating Distribution}
@@
 However, this peak diminishes when games involving unrated players are excluded, bringing the distribution closer to normality.
-This is illustrated in Figure~\ref{fig:all_games} and Figure~\ref{fig:rated_games}.
+This is illustrated in Figure~\ref{fig:transformation_plot}.
 Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the extremes.

### 2024-09-29 src/transform.py

@@ -10,8 +10,6 @@ def transform(df):
     df = df.rename(columns={'last_move_at': 'year'})
     df = df.rename(columns={'created_at': 'time'})
     df = df.rename(columns={'increment_code': 'time_control'})
-    df = df[df['victory_status'] != 'resign']
-    df = df[df['victory_status'] != 'outoftime']
     df = df.reset_index()
     df = df.drop(columns=['index'])
     df = updateELO(df)

### 2024-09-29 src/trasformation_plot.py

@@ -7,17 +7,15 @@ def tr_plot(df1, df2):
     # Generate random data for four datasets
     data1 = df1['white_rating']
     data2 = df2['white_rating']
-    data3 = df1['turns']
-    data4 = df2['turns']
 
-    data = [data1, data2, data3, data4]
+    data = [data1, data2]
 
     # Set up the figure and axes
-    fig, axs = plt.subplots(2, 2, figsize=(10, 10))  # 2x2 grid of plots
+    fig, axs = plt.subplots(2, figsize=(10, 10))
 
     for i, ax in enumerate(axs.flatten()):
         # Plot the histogram
-        if i % 2 == 1:
+        if i % 2 == 0:
             color = 'b'
         else:
             color = 'g'
@@ -33,7 +31,7 @@ def tr_plot(df1, df2):
         ax.plot(x, p, 'k', linewidth=2)
         
         # Title with the fit parameters
-        title1 = 'player rating' if i < 2 else 'turns'
+        title1 = 'player rating'
         title2 = ' before transformation' if i % 2 == 0 else ' after transformation'
         ax.set_title(title1+title2)
 

### 2024-09-29 Article/Article.tex
Updated Article

@@
 
-After these trasformations, we are left with 5,827 games which we will now begin to analyze.
+After these trasformations, we are left with 16,155 games which we will now begin to analyze.
+
+\subsection{Rating Distribution}
+The distribution of player ratings in the dataset initially resembles a normal distribution with a peak around the mean.
+However, this peak diminishes when games involving unrated players are excluded, bringing the distribution closer to normality.
+This is illustrated in Figure~\ref{fig:transformation_plot}.
+Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the extremes.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{ratingqq.png}
+    \caption{Quantile - Quantile plot of player rating in relation to fitted normal distribution}
+    \label{fig:rating_qq}
+\end{figure}
+
+\subsubsection*{Kolmogorov-Smirnov (KS) Test for Normality}
+We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution.
+The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed.
+However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size.
+This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.
 
@@
 
-\subsection{Rating Distribution}
-The distribution of player ratings in the dataset initially resembles a normal distribution with a peak around the mean.
-However, this peak diminishes when games involving unrated players are excluded, bringing the distribution closer to normality.
-This is illustrated in Figure~\ref{fig:transformation_plot}.
-Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the extremes.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\linewidth]{ratingqq.png}
-    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
-    \label{fig:rating_qq}
-\end{figure}
-
-\subsubsection{Kolmogorov-Smirnov (KS) Test for Normality}
-We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution.
-The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed.
-However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size.
-This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.
-
 \subsection{Relations Between Features}
@@
 
-\subsubsection{Mann-Whitney U Test for Game Length}
+\subsubsection*{Mann-Whitney U Test for Game Length}
 Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players.

### 2024-09-29 Article/Article.tex
Updated Article

@@
 This is illustrated in Figure~\ref{fig:transformation_plot}.
-Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the extremes.
+Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the tails, which are exacerbated by the QQ plot's sensitivity to deviations at the tail.
 
@@
     \includegraphics[width=0.8\linewidth]{ratingqq.png}
-    \caption{Quantile - Quantile plot of player rating in relation to fitted normal distribution}
+    \caption{Quantile - Quantile plot of white player rating in relation to normal distribution}
     \label{fig:rating_qq}
@@
 This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.
+
+\subsection{Relations Between Features}
+Analyzing the dataset’s correlation matrix reveals linear relationships between certain features.
+For example, there is a weak positive correlation between the players' average rating and game length.
+There are also weak negative correlations between the number of turns and both the absolute difference in player ratings and the 'winner' variable, where 'winner' is numerized as 0 for draws and 1 for games with a winner.
+From these correlations, we can tentatively conclude that:
+\begin{itemize}
+    \item Games involving more skilled players tend to last longer.
+
+    \item Games with a larger skill gap between players may end faster.
+
+    \item Draws tend to draw out longer then desicive games.
+\end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{corr_mat.png}
+    \caption{Heat map of the correlation matrix}
+    \label{fig:corr_mat}
+\end{figure}
+
+\subsection{Connection Between Player Rating and Game Length}
+To investigate whether player ratings are associated with game length, we divided players into two groups based on their ratings: those with ratings higher than 1500 (high-rated) and those with ratings lower than 1500 (low-rated).
+The threshold of 1500 was chosen because it represents the starting rating on the Lichess platform.
+The empirical cumulative distribution function (ECDF) of game lengths for both groups (Figure~\ref{fig:ecdf_plot}) shows that high-rated players generally play longer games.
+
+\subsubsection*{Mann-Whitney U Test for Game Length}
+Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players.
+Our null hypothesis was that high-rated players do not play longer games than low-rated players on average.
+The test returned a p-value of \(2.7 \times 10^{-73}\), which allows us to reject the null hypothesis.
+Therefore, we conclude that player rating significantly influences game length, with higher-rated players tending to play longer games.
 
@@
 
-\subsection{Relations Between Features}
-Analyzing the dataset’s convolution matrix reveals linear relationships between certain features.
-For example, there is a strong correlation between the ratings of white and black players in a game, which aligns with the expectation that players of similar ratings tend to compete against each other.
-However, no significant linear relationship was observed between the winner and the ratings of either player.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\linewidth]{ratingecdf.png}
-    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
-    \label{fig:ecdf_plot}
-\end{figure}
-
-\subsection{Connection Between Player Rating and Game Length}
-To investigate whether player ratings are associated with game length, we divided players into two groups based on their ratings: those with ratings higher than 1500 (high-rated) and those with ratings lower than 1500 (low-rated).
-The threshold of 1500 was chosen because it represents the starting rating on the Lichess platform.
-The empirical cumulative distribution function (ECDF) of game lengths for both groups (Figure~\ref{fig:ecdf_plot}) shows that high-rated players generally play longer games.
-
-\subsubsection*{Mann-Whitney U Test for Game Length}
-Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players.
-Our null hypothesis was that high-rated players do not play longer games than low-rated players on average.
-The test returned a p-value of \(2.7 \times 10^{-73}\), which allows us to reject the null hypothesis.
-Therefore, we conclude that player rating significantly influences game length, with higher-rated players tending to play longer games.
-
-\subsection{Summary of Results}
-The analysis demonstrates a clear statistical relationship between player rating and game length.
-The Mann-Whitney U test confirmed that higher-rated players play longer games.
-This finding supports the hypothesis that player rating is a factor in game duration, with the distribution of ratings further informing our understanding of game dynamics.
-
 \newpage

### 2024-09-29 main.py

@@ -7,6 +7,8 @@ import seaborn as sns
 from src.transform import transform
 from src.trasformation_plot import tr_plot
 from src.save_plot import save_plot
+from src.qq_rating_plot import qq_plot
+from src.correlation import corr_plot
 
 save_dir = os.path.abspath('figures')
 
@@ -19,3 +21,12 @@ df_tr = transform(df)
 fig = tr_plot(df, df_tr)
 
 save_plot(fig, 'transformation_plot', save_dir)
+
+fig = qq_plot(df_tr['white_rating']) 
+
+save_plot(fig, 'ratingqq', save_dir)
+
+fig = corr_plot(df_tr) 
+
+save_plot(fig, 'corr_mat', save_dir)
+

### 2024-09-29 src/correlation.py

@@ -1,0 +1,17 @@
+import statsmodels.api as sm
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+def corr_plot(df):
+
+    fig, ax = plt.subplots()
+    df['winner'] = df['winner'].replace({'white' : 1, 'black' : 1, 'draw' : 0})
+    corr_matrix = df[['turns', 'winner', 'ave_rating', 'dif_rating']].corr()
+    plt.figure(figsize=(8, 6))
+    sns.heatmap(corr_matrix, annot=True, cmap='seismic', center=0)
+    
+    return fig
+    
+    

### 2024-09-29 src/qq_rating_plot.py

@@ -1,0 +1,16 @@
+import statsmodels.api as sm
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+def qq_plot(data):
+
+    # Create QQ plot using statsmodels
+    fig, ax = plt.subplots()
+    sm.qqplot(data, line='s')
+    plt.xlabel('Theoretical Quantiles')
+    plt.ylabel('Sample Quantiles')
+    
+    return fig
+    
+    

### 2024-09-29 src/transform.py

@@ -37,4 +37,4 @@ def ave(white_rating, black_rating):
     return (white_rating + black_rating) // 2
 
 def dif(white_rating, black_rating):
-    return (white_rating - black_rating)
+    return np.abs(white_rating - black_rating)

### 2024-09-29 Article/Article.tex
Updated Article

@@
 
+\section{Methods}
+\subsection{Quantile - Quantile Plot}
+
+\subsection{Kolmogorov-Smirnov Test for Normality}
+The Kolmogorov-Smirnov (KS) test for normality evaluates whether a sample comes from a normal distribution.
+It uses the empirical cumulative distribution function (ECDF), $F_n(x)$, of the sample and compares it to the cumulative distribution function (CDF) of the normal distribution, $F(x)$.
+The test statistic, $D_n$, is defined as:
+\[D_n = \sup_x |F_n(x) - F(x)|\]
+
+The null hypothesis $H_0$ assumes that the sample comes from the normal distribution, while the alternative hypothesis $H_1$ posits that it does not.
+
+    If $D_n$ exceeds a critical value (based on the sample size and significance level), we reject $H_0$, concluding the sample deviates significantly from normality.
+
+The KS test is sensitive to both location and shape differences, but less powerful in detecting deviations in the tails of the distribution when compared to other tests like Shapiro-Wilk or Anderson-Darling.
+
+\subsection{Correlation Matrix}
+A correlation matrix is a table that displays the pairwise correlation coefficients between multiple variables in a dataset.
+Each element in the matrix represents the degree of linear relationship between two variables, typically measured by Pearson's correlation coefficient:
+\[r_{XY} = \frac{Cov(X,Y)}{\sigma_X\sigma_Y}\]
+
+\subsection{Mann-Whitney U Test}
+
 \newpage

### 2024-09-29 Article/Article.tex
YESSgit add .

@@
     \includegraphics[width=0.8\linewidth]{gamma_fit.png}
-    \caption{Gamma distribution fitted to the 'Turns' data}
+    \caption{Gamma distribution estimation of Turns}
     \label{fig:gamma_fit}
@@
 
-To better understand the distribution of the 'Turns' variable, we attempted to fit it to several probability distribution families.
-The Gamma distribution was selected first due to its ability to model skewed data.
-However, as shown in Figure \ref{fig:gamma_fit}, the Gamma distribution is too broad for this dataset, suggesting it does not fully capture the underlying structure of 'Turns.'
+Finally, the most complete way to gain knowlage on a dataset is to classify its distribution. To do that, let us look at the moment's information for Turns:
+\[Mean: 61.96255029402662\]
+\[Standard Deviation: 33.731730929355365\]
+\[Skewness: 0.9052618256476783\]
+\[Kurtosis: 1.4921324651335777\]
 
-To refine the fit, we combined the Gamma distribution with a Poisson distribution, which has a narrower shape.
-This combination results in a more accurate fit to the empirical data, as shown in Figure \ref{fig:gam_poi_fit}.
-The Gamma-Poisson mixture balances the excess spread of the Gamma distribution with the narrowness of the Poisson, providing a better representation of the 'Turns' distribution.
+The Kurtosis is too large for a normal distribution and too small for poisson.
+The skewness suggests an a-symetric distribution.A good general distribution family is the Gamma distribution family, but as seen in Figure~\ref{fig:gamma_fit} it lacks flexibility of shape.
+To gain more flexibility we have used the general gamma distribution family.
+But this won't add to our lack of Kurtosis.
+To do that we decided to average the general gamma with a poisson distribution and to fit them together.
+The results are shown in Figure~\ref{fig:gam_poi_fit}.
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{gam_poi_fit.png}
-    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
+    \includegraphics[width=0.8\linewidth]{gamma_poisson_fit.png}
+    \caption{Gamma-Poisson distribution estimation of Turns}
     \label{fig:gam_poi_fit}
@@
 \subsection{Mann-Whitney U Test}
+The \textbf{Mann-Whitney U test} is a non-parametric test used to determine whether there is a significant difference between two independent samples.
+The null hypothesis (\(H_0\)) assumes that the two distributions are identical, and the alternative hypothesis (\(H_a\)) assumes that one distribution tends to have larger or smaller values than the other.
+
+Procedure:
+
+Let \( n_1 \) and \( n_2 \) be the sample sizes of two independent groups, \( X_1, X_2, \dots, X_{n_1} \) and \( Y_1, Y_2, \dots, Y_{n_2} \).
+
+1. Rank all the observations from both groups together, assigning rank 1 to the smallest value, rank 2 to the second smallest, and so on.
+
+2. Compute the sum of the ranks for the first group \( R_1 \).
+
+3. The U statistic for the first group is given by:
+\[
+U_1 = n_1 n_2 + \frac{n_1(n_1 + 1)}{2} - R_1
+\]
+Similarly, compute the U statistic for the second group:
+\[
+U_2 = n_1 n_2 + \frac{n_2(n_2 + 1)}{2} - R_2
+\]
+
+4. The Mann-Whitney U statistic is the smaller of \( U_1 \) and \( U_2 \):
+\[
+U = \min(U_1, U_2)
+\]
+
+Normal Approximation (for large samples):
+
+When the sample sizes are large (\(n_1 > 20\) and \(n_2 > 20\)), the distribution of the U statistic can be approximated by a normal distribution with the following mean and standard deviation:
+
+\[
+\mu_U = \frac{n_1 n_2}{2}
+\]
+\[
+\sigma_U = \sqrt{\frac{n_1 n_2 (n_1 + n_2 + 1)}{12}}
+\]
+
+The standard normal \( Z \)-score is then computed as:
+\[
+Z = \frac{U - \mu_U}{\sigma_U}
+\]
+
+Decision Rule:
+
+To test the null hypothesis, compare the calculated U statistic (or Z-score for large samples) to the critical value from the Mann-Whitney U distribution.
+If the observed value of \( U \) (or \( Z \)) is extreme enough, reject \( H_0 \) in favor of \( H_a \).
 

### 2024-09-29 main.py

@@ -9,6 +9,8 @@ from src.trasformation_plot import tr_plot
 from src.save_plot import save_plot
 from src.qq_rating_plot import qq_plot
 from src.correlation import corr_plot
+from src.gamma_fit import fit_gamma
+from src.gamma_poisson_fit import fit_gamma_poisson
 
 save_dir = os.path.abspath('figures')
 
@@ -30,3 +32,10 @@ fig = corr_plot(df_tr)
 
 save_plot(fig, 'corr_mat', save_dir)
 
+fig = fit_gamma(df['turns'])
+
+save_plot(fig, 'gamma_fit', save_dir)
+
+fig = fit_gamma_poisson(df['turns'])
+
+save_plot(fig, 'gamma_poisson_fit', save_dir)

### 2024-09-29 src/gamma_fit.py

@@ -1,0 +1,28 @@
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+def fit_gamma(data):
+    shape, loc, scale = stats.gamma.fit(data, floc=0)  # Force location to 0 if appropriate
+
+    # Set up the figure and axes
+    fig, axs = plt.subplots()
+
+    # Generate a range of values for plotting the fitted distribution
+    x = np.linspace(1, 250, 250)
+
+    # Plot the histogram of the actual data
+    sns.histplot(data, kde=False, stat='density', label='Data')
+
+
+    pdf_fitted = stats.gamma.pdf(x, shape, loc=loc, scale=scale)
+    plt.plot(x, pdf_fitted, label='Fitted Gamma PDF', color='red')
+
+    # Add plot labels
+    plt.xlabel('Turns')
+    plt.ylabel('Density')
+    plt.legend()
+
+    return fig

### 2024-09-29 src/gamma_poisson_fit.py

@@ -1,0 +1,63 @@
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+import matplotlib.pyplot as plt
+import seaborn as sns
+from scipy.optimize import minimize
+
+def log_likelihood_gamma_poisson(params, data):
+    """
+    Calculates the log-likelihood of a mixture of Gamma and Poisson distributions.
+
+    Args:
+        params: A tuple containing the parameters: (a, c, loc, scale, lam, w).
+                a, c, loc, scale are the Gamma distribution parameters.
+                lam is the Poisson distribution parameter.
+                w is the weight of the Gamma distribution (Poisson weight is 1-w).
+        data: The observed data.
+
+    Returns:
+        The negative log-likelihood (for minimization).
+    """
+    a, c, loc, scale, lam, w = params
+    if a <= 0 or c <= 0 or scale <= 0 or lam <= 0 or w < 0 or w > 1:  # Parameter constraints
+        return np.inf
+
+    gamma_pdf = stats.gengamma.pdf(data, a, c, loc=loc, scale=scale)
+    poisson_pmf = stats.poisson.pmf(data, lam)
+
+    likelihood = w * gamma_pdf + (1 - w) * poisson_pmf
+    log_likelihood = np.sum(np.log(likelihood)) 
+
+    return -log_likelihood # Negative for minimization
+
+def fit_gamma_poisson(data):
+    # Initial guesses for parameters
+    initial_params = (2, 2, 10, 5, 25, 0.6)  # Example initial guess
+
+    # Set up the figure and axes
+    fig, axs = plt.subplots()
+
+    # Optimization (find maximum likelihood estimates)
+    result = minimize(log_likelihood_gamma_poisson, initial_params, args=(data,), method='Nelder-Mead')
+
+    # Fitted parameters
+    a_fitted, c_fitted, loc_fitted, scale_fitted, lam_fitted, w_fitted = result.x
+
+    print("Fitted parameters:")
+    print(f"  a: {a_fitted}, c: {c_fitted}, loc: {loc_fitted}, scale: {scale_fitted}")
+    print(f"  lambda: {lam_fitted}")
+    print(f"  w (Gamma weight): {w_fitted}")
+
+
+    # Plotting
+    x = np.linspace(1, data.max(), data.max()) # Ensure integer spacing
[…]

### 2024-09-29 Article/Article.tex
Updated Article

@@
 
-Finally, the most complete way to gain knowlage on a dataset is to classify its distribution. To do that, let us look at the moment's information for Turns:
+The most comprehensive way to analyze a dataset is through the classification of its distribution. To achieve this, we examine the moments of the distribution for the variable "Turns" as follows:
 \[Mean: 61.96255029402662\]
@@
 
-The Kurtosis is too large for a normal distribution and too small for poisson.
-The skewness suggests an a-symetric distribution.A good general distribution family is the Gamma distribution family, but as seen in Figure~\ref{fig:gamma_fit} it lacks flexibility of shape.
-To gain more flexibility we have used the general gamma distribution family.
-But this won't add to our lack of Kurtosis.
-To do that we decided to average the general gamma with a poisson distribution and to fit them together.
-The results are shown in Figure~\ref{fig:gam_poi_fit}.
+The kurtosis value (1.5) is higher than expected for a normal distribution, but lower than that of a Poisson distribution, indicating that neither distribution adequately fits the data.
+The positive skewness (0.9) implies an asymmetric distribution, with a longer tail on the right side.
+
+Given these characteristics, a natural candidate for modeling the data is the Gamma distribution family, known for handling positively skewed data.
+However, as demonstrated in Figure~\ref{fig:gamma_fit}, the Gamma distribution lacks sufficient flexibility to capture the dataset’s specific shape.
+To address this limitation, we explored the more general Generalized Gamma distribution, which offers greater shape flexibility.
+
+Despite this improvement, the Generalized Gamma distribution does not sufficiently account for the excess kurtosis.
+Therefore, we combined it with a Poisson distribution to create a mixture model that captures both the skewness and kurtosis properties.
+The results of this hybrid model are illustrated in Figure~\ref{fig:gam_poi_fit}.
 

### 2024-09-30 Article/Article.tex
Added statistical tests

@@
 The threshold of 1500 was chosen because it represents the starting rating on the Lichess platform.
-The empirical cumulative distribution function (ECDF) of game lengths for both groups (Figure~\ref{fig:ecdf_plot}) shows that high-rated players generally play longer games.
-
 \subsubsection*{Mann-Whitney U Test for Game Length}
 Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players.
-Our null hypothesis was that high-rated players do not play longer games than low-rated players on average.
+Our null hypothesis was that high-rated players' game length mean wasn't higher than low-rated players' mean.
 The test returned a p-value of \(2.7 \times 10^{-73}\), which allows us to reject the null hypothesis.

### 2024-09-30 main.py

@@ -11,6 +11,10 @@ from src.qq_rating_plot import qq_plot
 from src.correlation import corr_plot
 from src.gamma_fit import fit_gamma
 from src.gamma_poisson_fit import fit_gamma_poisson
+from src.KS_two_sample import KS_2test
+from src.dataset_statistics import dataset_statistics
+from src.KS_normal_test import KS_norm_test
+from src.MWU_test import MWU_test
 
 save_dir = os.path.abspath('figures')
 
@@ -28,14 +32,32 @@ fig = qq_plot(df_tr['white_rating'])
 
 save_plot(fig, 'ratingqq', save_dir)
 
+KS_norm_test(df_tr['ave_rating'].astype(int))
+
+#MWU_test()
+
 fig = corr_plot(df_tr) 
 
 save_plot(fig, 'corr_mat', save_dir)
 
-fig = fit_gamma(df['turns'])
+fig = fit_gamma(df_tr['turns'])
 
 save_plot(fig, 'gamma_fit', save_dir)
 
-fig = fit_gamma_poisson(df['turns'])
+fig, a_fitted, c_fitted, loc_fitted, scale_fitted, lam_fitted, w_fitted = fit_gamma_poisson(df_tr['turns'])
 
 save_plot(fig, 'gamma_poisson_fit', save_dir)
+
+dataset_statistics(df_tr['turns'])
+
+
+data = df_tr['turns']
+n = len(data)
+# Use a weighted random choice to sample from the mixture
+samples = np.zeros(n)
+for i in range(n):
+    if np.random.rand() < w_fitted:
+        samples[i] = stats.gengamma.rvs(a_fitted, c_fitted, loc=loc_fitted, scale=scale_fitted)
+    else:
+        samples[i] = stats.poisson.rvs(lam_fitted)
+KS_2test(data, samples)

### 2024-09-30 src/KS_normal_test.py

@@ -1,0 +1,9 @@
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+
+def KS_norm_test(data):
+    
+    ks_stat, ks_p_value = stats.kstest(data, 'norm', args=(np.mean(data), np.std(data)))
+
+    print(f"KS Statistic: {ks_stat}, p-value: {ks_p_value}")

### 2024-09-30 src/KS_two_sample.py

@@ -1,0 +1,12 @@
+from scipy.stats import kstest
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+
+def KS_2test(data1, data2):
+    # Perform the KS test
+    ks_statistic, p_value = stats.ks_2samp(data1, data2)
+
+    print("\nKS Test Results:")
+    print(f"  KS statistic: {ks_statistic}")
+    print(f"  p-value: {p_value}")

### 2024-09-30 src/MWU_test.py

@@ -1,0 +1,9 @@
+import scipy.stats as stats
+import numpy as np
+import pandas as pd
+
+def MWU_test(data1, data2):
+    statistic, p_value = stats.mannwhitneyu(data1, data2, alternative='greater')
+
+    print(f"Mann-Whitney U statistic: {statistic}")
+    print(f"P-value: {p_value}")

### 2024-09-30 src/dataset_statistics.py

@@ -1,0 +1,16 @@
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+
+def dataset_statistics(data):
+    # Calculate statistics
+    mean = np.mean(data)
+    std = np.std(data)
+    skew = stats.skew(data)
+    kurt = stats.kurtosis(data)
+
+    # Print statistics
+    print("Mean:", mean)
+    print("Standard Deviation:", std)
+    print("Skewness:", skew)
+    print("Kurtosis:", kurt)

### 2024-09-30 src/gamma_poisson_fit.py

@@ -60,4 +60,4 @@ def fit_gamma_poisson(data):
     plt.ylabel('Density')
     plt.legend()
 
-    return fig
+    return fig, a_fitted, c_fitted, loc_fitted, scale_fitted, lam_fitted, w_fitted

### 2024-09-30 Article/Article.tex
Updated Article

@@
 
-\subsection{Kolmogorov-Smirnov Test for Normality}
+\subsection{Kolmogorov-Smirnov Test}
 The Kolmogorov-Smirnov (KS) test for normality evaluates whether a sample comes from a normal distribution.
@@
 The KS test is sensitive to both location and shape differences, but less powerful in detecting deviations in the tails of the distribution when compared to other tests like Shapiro-Wilk or Anderson-Darling.
-
+\subsubsection*{2 sample KS test}
+Follows the same principles, except it compares two ECDF's from two samples.
 \subsection{Correlation Matrix}
@@
 
+\subsection{Gamma-Poisson distribution}
+
+\subsection{Nelder - Mead optimization method}
+
+\section{Discussion}
+
 \newpage

### 2024-09-30 Article/Article.tex
Updated Article

@@
 \subsection{Gamma-Poisson distribution}
+The GP distribution is a simple average between a general Gamma distribution(GGD) and a Poisson distribution weighted by a parameter w.
+\[f_{GP}(x|\alpha, \beta, \gamma, \lambda, w) = wf_{GGD}(x|\alpha, \beta, \gamma) + (1-w)f_P(\lambda)\]
+Let us find the MLE of GP:
+\[L_{GP}(\alpha, \beta, \gamma, \lambda, w) = \prod_{i=1}^n f_{GP}(x_i|\alpha, \beta, \gamma, \lambda, w)\]
+\[= \prod_{i=1}^n [wf_{GGD}(x|\alpha, \beta, \gamma) + (1-w)f_P(\lambda)]\]
+Now, by taking the negative log of the likelihood function, we can use minimizing optimization algorithms to find the optimal parameters numerically, since here an analytical solution is difficult to find.
 

### 2024-09-30 Article/Article.tex
Updated src

@@
 \subsection{Nelder - Mead optimization method}
+The Nelder-Mead optimization method is a direct search algorithm used for solving unconstrained nonlinear optimization problems without requiring gradient information.
+It operates on a simplex, a geometric shape consisting of $n+1$ vertices in an $n$-dimensional space.
+The algorithm iteratively updates this simplex by reflecting, expanding, contracting, or shrinking its vertices to explore the search space.
+These transformations aim to minimize an objective function $f(x)$ through comparisons of function values at the simplex vertices.
+Despite its simplicity, the method is known for efficiency in practice but does not guarantee convergence to a global minimum, especially in higher-dimensional or highly non-convex landscapes.
+It is widely applied in scenarios where gradient information is unavailable or expensive to compute.
 

### 2024-09-30 main.py

@@ -34,7 +34,16 @@ save_plot(fig, 'ratingqq', save_dir)
 
 KS_norm_test(df_tr['ave_rating'].astype(int))
 
-#MWU_test()
+# Classifying players into groups based on skill
+df_cl = df_tr.copy()
+def skill_classification(x):
+    if x < 1500:
+        return 'low'
+    else:
+        return 'high'
+
+df_cl['ave_rating'] = df_cl['ave_rating'].apply(skill_classification)
+MWU_test(df_cl[df_cl['ave_rating'] == 'high']['turns'], df_cl[df_cl['ave_rating'] == 'low']['turns'])
 
 fig = corr_plot(df_tr) 
 

### 2024-09-30 src/MWU_test.py

@@ -5,5 +5,7 @@ import pandas as pd
 def MWU_test(data1, data2):
     statistic, p_value = stats.mannwhitneyu(data1, data2, alternative='greater')
 
+    data1 = data1.dropna()
+    data2 = data2.dropna()
     print(f"Mann-Whitney U statistic: {statistic}")
     print(f"P-value: {p_value}")

### 2024-09-30 main.py
Updated src

@@ -70,3 +70,33 @@ for i in range(n):
     else:
         samples[i] = stats.poisson.rvs(lam_fitted)
 KS_2test(data, samples)
+'''
+fig, a_fitted, c_fitted, loc_fitted, scale_fitted, lam_fitted, w_fitted = fit_gamma_poisson(df_cl[df_cl['ave_rating'] == 'low']['turns'])
+
+save_plot(fig, 'gamma_poisson_fit_low', save_dir)
+
+data = df_cl[df_cl['ave_rating'] == 'low']['turns']
+n = len(data)
+# Use a weighted random choice to sample from the mixture
+samples = np.zeros(n)
+for i in range(n):
+    if np.random.rand() < w_fitted:
+        samples[i] = stats.gengamma.rvs(a_fitted, c_fitted, loc=loc_fitted, scale=scale_fitted)
+    else:
+        samples[i] = stats.poisson.rvs(lam_fitted)
+KS_2test(data, samples)
+
+fig, a_fitted, c_fitted, loc_fitted, scale_fitted, lam_fitted, w_fitted = fit_gamma_poisson(df_cl[df_cl['ave_rating'] == 'high']['turns'])
+
+save_plot(fig, 'gamma_poisson_fit_high', save_dir)
+
+data = df_cl[df_cl['ave_rating'] == 'high']['turns']
+n = len(data)
+# Use a weighted random choice to sample from the mixture
+samples = np.zeros(n)
+for i in range(n):
+    if np.random.rand() < w_fitted:
+        samples[i] = stats.gengamma.rvs(a_fitted, c_fitted, loc=loc_fitted, scale=scale_fitted)
+    else:
+        samples[i] = stats.poisson.rvs(lam_fitted)
+KS_2test(data, samples) '''

### 2024-09-30 src/KS_two_sample.py

@@ -1,4 +1,3 @@
-from scipy.stats import kstest
 import numpy as np
 import pandas as pd
 import scipy.stats as stats
@@ -7,6 +6,6 @@ def KS_2test(data1, data2):
     # Perform the KS test
     ks_statistic, p_value = stats.ks_2samp(data1, data2)
 
-    print("\nKS Test Results:")
+    print("\n2 Sample KS Test Results:")
     print(f"  KS statistic: {ks_statistic}")
     print(f"  p-value: {p_value}")

### 2024-09-30 Article/Article.tex
Updated src
Updated Article

@@
 \maketitle
+
+\begin{abstract}
+This study explores the relationship between \textbf{player ratings} and \textbf{chess game duration} using a dataset of online games.
+We aim to determine whether \textbf{higher-rated players} tend to play longer or shorter games and investigate other factors influencing game length, such as \textbf{time controls} and \textbf{game outcomes}.
+A cleaned dataset was analyzed through statistical techniques, including the \textbf{Kolmogorov-Smirnov test} for normality, the \textbf{Mann-Whitney U test} for differences in game length between rating groups, and \textbf{correlation analysis}.
+Results indicate that \textbf{higher-rated players} generally engage in longer games, while greater disparities in player ratings lead to shorter matches.
+Additionally, \textbf{draws} tend to prolong game duration compared to decisive outcomes.
+The game's turn count distribution shows positive skewness and excess kurtosis, leading us to model it with a \textbf{Gamma-Poisson mixture}.
+This model effectively captures the underlying distribution characteristics, providing further insights into the statistical patterns governing chess match dynamics.
+\end{abstract}
 
@@
 The most comprehensive way to analyze a dataset is through the classification of its distribution. To achieve this, we examine the moments of the distribution for the variable "Turns" as follows:
-\[Mean: 61.96255029402662\]
-\[Standard Deviation: 33.731730929355365\]
-\[Skewness: 0.9052618256476783\]
-\[Kurtosis: 1.4921324651335777\]
+\[Mean: 61.962\]
+\[Standard Deviation: 33.731\]
+\[Skewness: 0.9052\]
+\[Kurtosis: 1.492\]
 
@@
 The results of this hybrid model are illustrated in Figure~\ref{fig:gam_poi_fit}.
+This fit actually performs amazingly for the 2 sample KS test, even sometimes beating it, with a statistic of $0.01429$ and a p value of $0.073$, which is above the arbitrary threshold of 0.05.
 

### 2024-09-30 Article/Article.tex
updated article

@@
 
+\subsection{The Effect of the Time Controls on Game Length}
+To check if the game's time controls have a significant correlation with turns, we checked a 2 sample KS test for the turns in rapid and Classical, which represent the vast majority of the games in the dataset.
+The resulting p value was $0.322$, which means the Null hypothesis of the two datasets being the same stands.
+From this we can conclude that there is no significant correlation between Time Controls and Game Length.
+
 \begin{figure}[H]
@@
 It is widely applied in scenarios where gradient information is unavailable or expensive to compute.
+We
 

### 2024-09-30 main.py

@@ -49,6 +49,9 @@ fig = corr_plot(df_tr)
 
 save_plot(fig, 'corr_mat', save_dir)
 
+print("Time Control test")
+KS_2test(df_tr[df_tr['time_control'] == 'Classical']['turns'], df_tr[df_tr['time_control'] == 'Rapid']['turns'])
+
 fig = fit_gamma(df_tr['turns'])
 
 save_plot(fig, 'gamma_fit', save_dir)

### 2024-09-30 src/transform.py

@@ -16,6 +16,28 @@ def transform(df):
     df = df.rename(columns={'white_id': 'ave_rating'})
     df = df.rename(columns={'black_id': 'dif_rating'})
 
+    def chess_time_class(time_control: str) -> str:
+        # Split the time control into minutes (X) and increment (Y)
+        try:
+            X, Y = map(int, time_control.split('+'))
+        except ValueError:
+            return "Invalid time control format"
+        
+        # Convert X (minutes) to total seconds and add the increment per move
+        total_time = X * 60 + Y * 40  # Assume 40 moves for standard calculation
+        
+        # Determine the time class based on total time
+        if total_time <= 180:
+            return "Bullet"
+        elif total_time <= 600:
+            return "Blitz"
+        elif total_time <= 1800:
+            return "Rapid"
+        else:
+            return "Classical"
+    
+    df['time_control'] = df['time_control'].apply(chess_time_class)
+
     for index, row in df.iterrows():
         df.at[index, 'ave_rating'] = ave(row['white_rating'], row['black_rating'])
         df.at[index, 'dif_rating'] = dif(row['white_rating'], row['black_rating'])  

### 2024-09-30 Article/Article.tex
updated tex file

@@
 \subsection{Quantile - Quantile Plot}
+Quntile-Quantile (QQ) plot is a tool used in order to determine the similarity between two distributions graphically.
+It is only a visual check over the plot, and does not provide an absolute detemination.
+In QQ plot, each one of the two distribution, might be samples of theory distribution, are divided into quantiles and is being ordered in an ascending order.
+Than, each pair of the same quantile is presented in the QQ plot as a point.
+The closer the dotted graph to a $y=x$ graph, the more similar the two distributions are.
+Because in perfect matching each quantile will be the same in bot distributions.
 
@@
 \section{Discussion}
+We did not reach a test which determine the players rating distribution as a normal distribution.
+Although we did get very close to it, be using part of the data.
+We are speculate that it should behave as a normal distribution, and there might be other reasonable filters on the data, in addition to refer only to rated games, or by collecting more data to the statistical analysis.
+
+Mann Whitney U test has being used because the distribution of the players rating did not determined.
+If a normal distribution, or any other known distribution, will be achieved by a test for the players rating, we might use some more specific test rather than Mann Whitney U test, such as t-test.
+
+In order to determine the correlation between players rating and game length we chose 1500 as a threshold for splitting the players to high and low rated.
+The choose of 1500 for a threshold between high and low rated players makes sense as it is the starting rate of any player, but it might be better to split the players into more groups in order to see actual correlation between the players rating and the number of turns in their games.
 

### 2024-09-30 Article/Article.tex
update bibliography, link to github and pdf

@@
 This model effectively captures the underlying distribution characteristics, providing further insights into the statistical patterns governing chess match dynamics.
+\newline \newline
+The code for this project is available at \href{https://github.com/nakardn/Chess-Statistical-Analysis}{https://github.com/nakardn/Chess-Statistical-Analysis}
 \end{abstract}
@@
 This is illustrated in Figure~\ref{fig:transformation_plot}.
-Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the tails, which are exacerbated by the QQ plot's sensitivity to deviations at the tail.
+Additionally, the QQ plot (Figure~\ref{fig:rating_qq}) supports the similarity between the player rating distribution and a normal distribution, albeit with deviations at the tails, which are exacerbated by the QQ plot's sensitivity to deviations at the tail \cite{qqplotwiki}.
 
